Gravitational field calculation in spherical coordinates using variable densities in depth

### 2019-04-04 manuscript/manuscript.tex
Add sinusoidal density function disclaimers (#53)
Fixes #46

\begin{summary}
We present a new methodology to compute the gravitational fields generated by
tesseroids (spherical prisms) whose density varies with depth according to
an arbitrary continuous function.
It approximates the gravitational fields through the Gauss-Legendre Quadrature along
with two discretization algorithms that automatically control its accuracy by adaptively
dividing the tesseroid into smaller ones.
The first one is a preexisting two dimensional adaptive discretization algorithm that
reduces the errors due to the distance between the tesseroid and the computation point.
The second is a new density-based discretization algorithm that
decreases the errors introduced by the variation of the density function with depth.
The amount of divisions made by each algorithm is indirectly controlled
by two parameters: the distance-size ratio and the delta ratio.
We have obtained analytical solutions for a spherical shell with radially variable
density and compared them to the results of the numerical model for linear,
exponential, and sinusoidal density functions.
These comparisons allowed us to obtain optimal values for the distance-size and
delta ratios that yield an accuracy of 0.1\% of the analytical solutions.
The resulting optimal values of distance-size ratio for the gravitational potential and
its gradient are 1 and 2.5, respectively.
The density-based discretization algorithm produces no discretizations in the linear
density case, but a delta ratio of 0.1 is needed for the exponential and the
sinusoidal density functions.
These values can be extrapolated to cover most common use cases.
However, the distance-size and delta ratios can be configured by the user to increase
the accuracy of the results at the expense of computational speed.
Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in
Argentina with a maximum depth of over 5000~m, using an exponential density function.
\end{summary}

@@

We empirically determined optimal values for $D$ and $\delta$ by comparing the numerical
results with analytical solutions.
Our analysis included a range of tesseroid models and computation grids as well as
common density functions.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimal values were a linear, an
exponential, and a sinusoidal function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
Through this analysis, we have obtained optimal values for the distance-size ratio $D$
of 1 and 2.5 for the potential and its gradient, respectively.
We analysed the error for exponential functions ranging from smooth to sharp and
sinusoidal function of different wavelengths to test the accuracy of the density-based
discretization.
Values of $\delta$ equal to 0.1 are sufficient to guarantee a 0.1\% accuracy for most
realist cases tested.
These results could be extrapolated to other continuous density functions that are
sufficiently smooth, as is the case for most common geophysical applications.
Smaller values of $\delta$ can be used for more highly variable density functions to
increase the accuracy of the integration accordingly.

>>>>

\begin{summary}
We present a new methodology to compute the gravitational fields generated by
tesseroids (spherical prisms) whose density varies with depth according to
an arbitrary continuous function.
It approximates the gravitational fields through the Gauss-Legendre Quadrature along
with two discretization algorithms that automatically control its accuracy by adaptively
dividing the tesseroid into smaller ones.
The first one is a preexisting two dimensional adaptive discretization algorithm that
reduces the errors due to the distance between the tesseroid and the computation point.
The second is a new density-based discretization algorithm that
decreases the errors introduced by the variation of the density function with depth.
The amount of divisions made by each algorithm is indirectly controlled
by two parameters: the distance-size ratio and the delta ratio.
We have obtained analytical solutions for a spherical shell with radially variable
density and compared them to the results of the numerical model for linear,
exponential, and sinusoidal density functions.
The heavily oscillating density functions are intended only to test the algorithm to its
limits and not to emulate a real world case.
These comparisons allowed us to obtain optimal values for the distance-size and
delta ratios that yield an accuracy of 0.1\% of the analytical solutions.
The resulting optimal values of distance-size ratio for the gravitational potential and
its gradient are 1 and 2.5, respectively.
The density-based discretization algorithm produces no discretizations in the linear
density case, but a delta ratio of 0.1 is needed for the exponential and most sinusoidal
density functions.
These values can be extrapolated to cover most common use cases, which are simpler than
oscillating density profiles.
However, the distance-size and delta ratios can be configured by the user to increase
the accuracy of the results at the expense of computational speed.
Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in
Argentina with a maximum depth of over 5000~m, using an exponential density function.
\end{summary}

@@

We empirically determined optimal values for $D$ and $\delta$ by comparing the numerical
results with analytical solutions.
Our analysis included a range of tesseroid models and computation grids as well as
common density functions.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimal values were a linear, an
exponential, and a sinusoidal function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
Through this analysis, we have obtained optimal values for the distance-size ratio $D$
of 1 and 2.5 for the potential and its gradient, respectively.
We analysed the error for exponential functions ranging from smooth to sharp and
sinusoidal function of different wavelengths to test the accuracy of the density-based
discretization.
It worth noting that the sinusoidal density tests were carried out in order to test the
algorithm to the extreme and not to emulate a real world scenario.
Values of $\delta$ equal to 0.1 are sufficient to guarantee a 0.1\% accuracy for most
realist cases tested.
These results could be extrapolated to other continuous density functions that are
sufficiently smooth, as is the case for most common geophysical applications.
Smaller values of $\delta$ can be used for more highly variable density functions to
increase the accuracy of the integration accordingly.
